### 2025-01-15 sections/2-methods.tex
complexity rewrite for sieve

We start with the assumption that the LFD near the query does not differ significantly from that of nearby clusters, and we consider leaf clusters with cardinalities near $k$.
Let $d$ be the LFD in this region.
Then the number of leaf-clusters in $Q$ is bounded above by $2d$, e.g.,\,we could have a cluster overlapping the query ball at each end of each of $\lceil d \rceil$ mutually-orthogonal axes.
In the worst-case scenario for tree-search, these leaf clusters would all come from different branches of the tree, and so tree-search looks at $2 \cdot \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X)$ clusters.
Thus, the asymptotic complexity is $T \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big)$.
For leaf-search, the output size and scaling factor are the same as in Repeated $\rho$-NN, and so the asymptotic complexity is $L \coloneqq \mathcal{O} \left( k \cdot \bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d \right)$.

The asymptotic complexity of Breadth-First Sieve is dominated by the QuickSelect algorithm to calculate $\tau$.
Since this method is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity is
\begin{equation}
    \mathcal{O} \Big( (T + L ) \log (T + L ) \Big).
    \label{eq:methods:breadth-first-sieve-complexity}
\end{equation}

For Depth-First Sieve, since we use two priority queues, the asymptotic complexity is dominated by the priority queue operations.
Thus, the complexity is
\begin{equation}
    \mathcal{O} \Big( T \log T + L \log k \Big).
    \label{eq:methods:depth-first-sieve-complexity}
\end{equation}

>>>>

\begin{theorem} 
    Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$. Then, for dataset $X$ and query $q$, the time complexity of performing Breadth-First Sieve search is \begin{equation}
        \mathcal{O} \Big( (T + L ) \log (T + L ) \Big)
        \label{eq:methods:breadth-first-sieve-complexity}
    \end{equation} and the the time complexity of performing Depth-First Sieve search is \begin{equation}
        \mathcal{O} \Big( T \log T + L \log k \Big).
        \label{eq:methods:depth-first-sieve-complexity}
    \end{equation}
\end{theorem}

\begin{proof}
Since $q$ is sampled from the same distribution as $X$, the LFD near $q$ should not differ significantly from the LFDs of clusters near $q$ at the scale of the distance from the query to the $k^{th}$ nearest neighbor.

Consider leaf clusters with cardinalities near $k$. Let $d$ be the LFD in this region, and $Q$ denote the set of candidate points and clusters.
Then the number of leaf-clusters in $Q$ is bounded above by $2d$, where the bound is achieved if we have a cluster overlapping the query ball at each end of each of $\lceil d \rceil$ mutually-orthogonal axes.
In the worst-case scenario for tree-search, these leaf clusters would all come from different branches of the tree, and so tree-search looks at $2 \cdot \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X)$ clusters.
Thus, the asymptotic complexity is $T \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big)$.
For leaf-search, the output size and scaling factor are the same as in Repeated $\rho$-NN, and so the asymptotic complexity is $L \coloneqq \mathcal{O} \left( k \cdot \bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d \right)$.

The asymptotic complexity of Breadth-First Sieve is dominated by the QuickSelect algorithm to calculate $\tau$.
Since this method is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity is
\begin{equation*}
    \mathcal{O} \Big( (T + L ) \log (T + L ) \Big),
\end{equation*} as claimed.

For Depth-First Sieve, since we use two priority queues, the asymptotic complexity is dominated by the priority queue operations.
Thus, the complexity is
\begin{equation*}
    \mathcal{O} \Big( T \log T + L \log k \Big), 
\end{equation*} as claimed. 
\end{proof}
